Generalized Tractability

### 2023-06-07 GeneralizedTractability_230529.tex
Updates from Overleaf

Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly $T$-separable tractability
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation*}
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = \lceil L_{(p,q)\psi(d,q)}\rceil}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\] for some $p,q >0$ and $L_{(p,q)}\in\naturals$.
\end{example}

\begin{example}[Non-separable tractability]
    Suppose we define the tractability function $T$ by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(\phi(\varepsilon^{-1})\, \psi(d))\}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp\in [0,\infty).
\]  with non-decreasing functions $\phi : (0,\infty) \to (0,\infty)$ and $\psi : \naturals \to (0,\infty)$ and $\vp = p$. 
We say that a problem is non-separable tractable with parameter $\vp$ iff there exists a positive $C_{\vp}$ such that

@@

Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for non-separable tractability
\[
\sup_{d\in\mathbb{N}}\frac{1}{\psi(d)}\sum_{i=Lp\,\psi(d)} \phi(\lambda_{i,d})\quad \text{for some } p \geq 0, L_p \in \naturals.
\]
We have the following special cases of non-separable tractability. \end{example}
\begin{example}[Algebraic quasi-polynomial tractability]

As defined in \cite{GW11}, we say that a problem is \emph{quasi-polynomially tractable (QPT)} if there exist two constants $C$, $t > 0$, such that
\begin{equation}\label{expqpt}
    \comp(\varepsilon,d) \leq C\exp\{t(1+\log(\varepsilon^{-1}))(1+\log(d))\}
\end{equation}
for all $\varepsilon \in (0,1]$ and $d\in\mathbb{N}$.\\

\noindent Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\,d^{-p}\sum_{i=Lp\,d^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)} < \infty \qquad \text{for some } p \geq 0, L_p \in \naturals.\]
\end{example}
\begin{example}[Exponential quasi-polynomial tractability]

@@

Suppose that the domain of interest $(\varepsilon^{-1},d)$ is not all of $(0,\infty) \times \naturals$ but some subset
\begin{equation}
    \Omega  := \{(\varepsilon^{-1},d) : (E(d),\infty), \ d \in \naturals \}, \qquad \text{where }E : \naturals \to [0,\infty).
\end{equation}
We then expect that the equivalent conditions for (strong) tractability will be similar in form, but weaker than, the conditions in Theorems  \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.

@@

For any $d\in\naturals$, let
\begin{align*}
L_{\vp'} & := \left[C_{\vp'} + \frac{2}{T(0,1,\vp')}   \right]
\intertext{and note that }
 \min(i(\infty,d,\vp'),2) & \le i(\infty,d,\vp') +1  = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +2 \le  \left[C_{\vp'} + \frac{2}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  L_{\vp'}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
This is the inequality needed for the lower limit of the sum.

@@

\item If $M(d) <\lceil L_{\vp'} T(0,d,\vp')\rceil$, then 
\[
\sum_{i = \lceil L_{\vp'} T(0,d,\vp')\rceil }^{M(d)} \frac{1}{T (\lambda_{i,d}^{-1},1, \vp')} = 0,
\]
which does not affect the finiteness of $S_{\vp'}$.

@@

\fred{Are there cases that we can discuss where we do not have (strong) tractability, but we do have (strong) tractability on a restricted domain, such as $(0,\infty) \times \{1, \ldots, d_{\max}] \cup (\varepsilon_{\min}^{-1},\infty) \times \naturals$}

>>>>

Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly $T$-separable tractability
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty.
\end{equation*}
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = \lceil L_{(p,q)}\psi(d,q)\rceil}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\] for some $p,q >0$ and $L_{(p,q)}\in\naturals$.
\end{example}

\begin{example}[Non-separable tractability]
    Suppose we define the tractability function $T$ by
\[
 T(\varepsilon^{-1},d,p)= \exp\{p \, \phi(\varepsilon^{-1})\, \psi(d)\} = [\exp\{ \phi(\varepsilon^{-1})\}]^{p \, \psi(d)}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p\in [0,\infty).
\]  with non-decreasing functions $\phi : (0,\infty) \to (0,\infty)$ and $\psi : \naturals \to (0,\infty)$. 
We say that a problem is non-separable tractable with parameter $\vp$ iff there exists a positive $C_{\vp}$ such that

@@

Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for non-separable tractability
\[
%\sup_{d\in\mathbb{N}}\frac{1}{\psi(d)}\sum_{i=Lp\,\psi(d)} \phi(\lambda_{i,d})\quad \text{for some } p \geq 0, L_p \in \naturals.
\sup_{d\in\mathbb{N}}\sum_{i=L_p\,\phi(0)^{p \, \psi(d)}} [\exp\{ \phi(\lambda_{i,d}^{-1})\}]^{p \, \psi(d)}\quad \text{for some } p \geq 0, L_p \in \naturals.
\]
\end{example}
For $\phi$ and $\psi$ of special form, the condition above can be simplified further as in the following two examples of non-separable tractability. 
\begin{example}[Algebraic quasi-polynomial tractability]

As adefined in \cite{GW11}, we say that a problem is \emph{quasi-polynomially tractable (QPT)} if there exist two constants $C$, $p > 0$, such that
\begin{equation}\label{expqpt}
    \comp(\varepsilon,d) \leq C\exp\{p(1+\log(\varepsilon^{-1}))(1+\log(d))\}
\end{equation}
for all $\varepsilon \in (0,1]$ and $d\in\mathbb{N}$.

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\] where in this case $\vp = p$. This can be rewritten as 
\[
T(\varepsilon^{-1},d,\vp) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\,d^{-p}\sum_{i=Lp\,d^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)} < \infty \qquad \text{for some } p \geq 0, L_p \in \naturals.\]
\end{example}
\begin{example}[Exponential quasi-polynomial tractability]

@@

Suppose that the domain of interest $(\varepsilon^{-1},d)$ is not all of $(0,\infty) \times \naturals$ but some subset
\begin{equation}
    \Omega  := \{(\varepsilon^{-1},d) : \varepsilon \in (E(d),\infty), \ d \in \naturals \}, \qquad \text{where }E : \naturals \to [0,\infty).
\end{equation}
We then expect that the equivalent conditions for (strong) tractability will be similar in form, but weaker than, the conditions in Theorems  \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.

@@

For any $d\in\naturals$, let
\begin{align*}
L_{\vp'} & := C_{\vp'} + \frac{2}{T(0,1,\vp')},
\intertext{and note that }
 \max(i(\infty,d,\vp'),2) & \le i(\infty,d,\vp') +1  = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +2 \le  \left[C_{\vp'} + \frac{2}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  L_{\vp'}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
This is the inequality needed for the lower limit of the sum.

@@

\item If $M(d) <\lceil L_{\vp'} T(0,d,\vp')\rceil$, then 
\[
\sum_{i = \lceil L_{\vp'} T(0,d,\vp')\rceil }^{M(d)} \frac{1}{T (\lambda_{i,d}^{-1},d, \vp')} = 0,
\]
which does not affect the finiteness of $S_{\vp'}$.

@@

\fred{Are there cases that we can discuss where we do not have (strong) tractability, but we do have (strong) tractability on a restricted domain, such as $(0,\infty) \times \{1, \ldots, d_{\max}] \cup (\varepsilon_{\min}^{-1},\infty) \times \naturals$?}
